Probing the Deuteron at Very Large Internal Momenta

### 2020-10-07 MY_PUBLICATIONS/PRL/FINAL_DRAFT/PRL_Submission/prl_response/prl_reply/reply_to_RefereeA.txt

--------------------------------
Reply to: Report of Referee A
--------------------------------

A concise, point-by-point response for each of the recommendations and criticisms made by referee A is shown below:

-------------
Referee A:
-------------
1) Universality: The nuclear-factorization assumption put forward by Ciofi and Frankfurt & Strikman, and more recently
developed to a quantitative approach using the generalized contact formalism, directly connect the high momentum
measurement of the deuteron to all atomic nuclei and nuclear systems. This universality steam from the dominance of
SRC pairs in the high momentum tail of nuclear wave-function. A discussion of this connection, with proper reference
to the theoretical works mention above, and to the experimental works that show the important of np-SRC pairs with the
deuteron quantum number (Subedi et al., Korover et al., Duer et al., etc.), is missing.
----------
Response:
----------
The introductory (1st) paragraph was re-written in accordance with the suggestions made above.

-------------
Referee A:
-------------
2) Relativity: The current manuscript does not discuss relativistic effects even though one could expect them to be
significant in the measured momenta range. Questions that immediately come to mind include to what extend the new data
supports a non-relativistic description of the deuteron wave function relevant for the high momenta measured here? If
they do, why? What can one learn about how to properly account for relativistic corrections and relativity in general
for nuclear systems? Do the new data constrain our limited ability to describe the relativistic aspects of nuclear structure
at short distance / high momenta? For example, Schmidt et al. [Nature 578 (2020) 540] recently utilized the light-cone formalism
of Frankfurt & Strikman to gain quantitative insight into these issues. The formalism they follow seem to be quite simple, perhaps
too simplistic, but still provide valuable insight. Can the authors of the present work study the impact of relativistic corrections
to the PWIA cross-section in a similar manner?
----------
Response:
----------
Our analysis demonstrated that there is no simplistic way of estimating relativistic effects since even in the kinematics
in which FSI are small they are still comparable with the expected relativistic effects. The current calculations are based on
the virtual nucleon approximation in which  relativistic effects are accounted for in the propagator of the bound nucleon, for the
vertex of electron-bound nucleon interaction as well as for FSI. To switch off relativistic effects means to remove all these effects
which will return us to the standard Glauber theory.  But in our previous paper [Phys.Rev.Lett. 107, 262501 (2011)],  we already
demonstrated that Glauber theory will not describe the main feature of the reaction. For example at 400 MeV/c it will predict maximum FSI
to occur at 90 degrees of the spectator angle, while the experiment clearly shows that maximum FSI is at 70 degrees.

The current limitation of the virtual nucleon approximation is that it still uses a non-relativistic deuteron
wave function [M. Sargsian, Private communication, Oct. 2020] and assumes that vacuum fluctuation effects are not significant.
It is expected that the latter will grow with the increase of the recoil momentum of the spectator nucleon. 
One approach that addresses this issue is the Light-Front approximation, however, the implementation of this approximation 
requires a consistent calculation of all the above discussed factors (i.e. propagator, vertex, FSI and wave function).
Hence, relativistic effects cannot simply be "turned off", since they contribute in different components of 
the scattering amplitude.

Light-front calculations are currently underway [M.Sargsian, Private communication, Oct. 2020) which we hope will be used 
to compare with our data for the large paper.

-------------
Referee A:
-------------
3) Nucleon-Nucleon interaction: Here the authors make the connection clear. However no modern EFT calculations are compared
with the new data. It is true that these are limited by strong regulators are high-momenta, however there are several interactions
such as the Norfolk potentials [M. Piarulli] that do suffer letter from such effects. In addition, even with high-momentum
regulators, it is interesting to observe out to what momenta can these interactions be coupled with a high- resolution one-body
interaction model and still explain experimental data. Recent analysis of nuclear form-factor data suggest this can be quite high.

In addition, there is no reference to relevant previous works that study nuclei heavier than deuterium. These include
A=3 [124 (2020) 212501; PRL 94 (2005) 082305; 94 (2005) 192302] and 12 [Nature 578 (2020) 540] that study the NN force
to short distance structure of SRC deuteron like pairs.
----------
Response:
----------
Similar to the above comment (2) regarding relativistic calculations to account for the effects of different
wave functions, it requires a complete calculation including the FSI effects, which is out of scope of the current paper. 
We hope that the theory groups who are collaborating with us will be able to provide such calculations which will be
presented in a larger paper. These studies are currently in progress.

-------------
Referee A:
-------------
4) More general comment: Not enough information is given on the measurements. The data was taken using a new detection system
(SHMS at Hall C in 12 GeV setting) and the reader is referred only to unpublished thesis (with no Web link) for details.
In addition, the analysis itself is not transparent enough to the reader. More information should be provided in the form of
supplementary materials or similar. See details below.
----------
Response:
----------
I have published my thesis on arXiv and reference it in the PRL manuscript. I also added supplementary materials.

-------------
Referee A:
-------------
5) The scale of Fig. 2 is misleading as it is very hard to assess the quality of the agreement between the data and
theory in the 0 – 700 MeV/c region. Please add an insert with a blowup on that region.
----------
Response:
----------
I have added an inset with a blowup for Fig. 2, subplots (a) and (b).

-------------
Referee A:
-------------
6) I noticed (and appreciate) the availability of the crosse-section data in txt format and the same should be done
for the relevant calculations so that the interested reader can examine the results in more detail.
----------
Response:
----------
I have added additional raw data files with the theoretical calculations as part of the
supplemental material.

-------------
Referee A:
-------------
7) Page 1 introduction: Should place more emphasis on the importance to the broader physics community, as suggested above.
Elucidating the short-distance structure of nuclei and nuclear force via deuteron measurements is a fascinating problem of
a broad interest. However, the present work does not, in my opinion, meet this criterion.
----------
Response:
----------
As mentioned in response to (1), I re-wrote the introductory paragraph.

-------------
Referee A:
-------------
8) Page 2 introduction "determining whether or to what extent the description of nuclei in terms of nucleon/meson degrees of
freedom is still valid before having to include explicit quark degrees of freedom”. Do the authors really think that (e,e’p)
measurement can address the question of nucleon/meson and partonic degrees of freedom? At the end of the day the data are compared
with calculations that have large freedom in their description of the off-shell electron-nucleon cross-section and even more so in
the modeling of the short-distance part of the NN interaction. Therefore, even if partonic physics becomes important at high-momenta,
it can probably be masked by effective nucleon- based interaction that will fit the data. In any case, it is ok if the authors want
to make such connections, however, it will probably be better to do so as part of a broader discussion of the relation between SRC
studies and the partonic structure of nuclei, e.g. by referencing the EMC effect and the EMC-SRC connections published in the last years.
----------
Response:
----------
This part the referee refers to was removed in the introductory paragraph.

-------------
Referee A:
-------------
9) The distributions of the fundamental measured quantities such as Q2, xB , ω, Pp should be shown. The measurements were done in 3 momentum bins.
The overlap and the matching between the bins should be shown. This can be a part of a supplementary materials document that allow the referees
and the reader to assess the quality of the data.
----------
Response:
----------
I have added Q2 distribution (for 80 MeV/c setting) in the Supplemental Material. I have also
added the missing momentum distributions for each of the central settings (80, 580 and 750 MeV/c) *** Maybe I can add Q2, xB, w and Pp for each setting (smaller plots)

-------------
Referee A:
-------------
10) How the was the contribution of the target walls removed / subtracted?
----------
Response:
----------
The background from target cell wall was found to be negligible at the deuteron kinematics
(See Supplemental Material)

-------------
Referee A:
-------------
11) “hits in three of the four scintillator planes.” At least three?
----------
Response:
----------
Yes, at least 3/4 planes. I have clarified this in the manuscript.

-------------
Referee A:
-------------
12) What about systematic uncertainties due to software cuts, acceptance corrections etc.?
----------
Response:
----------
See Supplemental Materials.

-------------
Referee A:
-------------
13) What is the size of the radiation correction? Please detail on its calculation procedure.
----------
Response:
----------
See Supplemental Materials.

-------------
Referee A:
-------------
14) “a weighted average of the cross sections were taken in the overlapping regions of pr”. Show consistency. 
----------
Response:
----------
To clarify, a weighted average of the reduced cross sections were taken for overlapping regions in pr. See example
of overlapping pr bins in Fig. 14 of Supplemental Materials.

-------------
Referee A:
-------------
15) “cach” should be “each”.
----------
Response:
----------
I did not find this typo in the manuscript.

-------------
Referee A:
-------------
16) Fig. 2 and relevant discussion on Page 5: what about EFT calculations below and above their cutoffs?

Pages 5-6 include extensive phenomenological verbal description of Fig 2. However, what it misses a discussion
of the possible meaning of what is actually observed? For example: “the MS CD-Bonn model was able to partially
describe the data over a larger range...” – what does it mean? “all models were unable to describe the data”
above some momentum-what that means?
----------
Response:
----------
The comparison between the measured (reduced) cross sections and the currently available
complete and state of the art calculations illustrates the limit to which a non-relativistic
wave function that is the solution of the Schroedinger equation is valid and able to describe
experimental data that probe the high momentum region of the np system in the most direct way
possible.

As was mentioned before, the theoretical interpretation of these results is
out of the scope of this paper and is currently being investigated by the theory
groups we are collaborating with. Our goal was to simply present these new and
interesting results.

### 2020-10-07 MY_PUBLICATIONS/PRL/FINAL_DRAFT/PRL_Submission/prl_response/prl_reply/MARKED/ms_revised.tex
added PRL reply

practically unexplored and not well understood. At such small inter-nucleon distances, the nucleon-nucleon ($NN$) interaction is expected to become repulsive and the interacting
nucleons begin to overlap. The dynamics in this short distance region are directly related to the dynamics of two-nucleon short range correlations (SRC) observed in $A>2$ nuclei \cite{PhysRevC.68.014313,PhysRevLett.96.082501,PhysRevLett.99.072501,Fomin_2017,Hen2014,RevModPhys.89.045002}.
Short-range studies of the deuteron are also important in determining whether, or to what extent, the description of nuclei in terms of nucleon/meson degrees of freedom is still valid, before
having to include explicit quark degrees of freedom, an issue of fundamental importance in nuclear physics \cite{sargsian_2015}. As of the present time, there are only a few nuclear physics experiments for
which a transition between nucleonic to quark degrees of freedom has been observed \cite{PhysRevLett.81.4576,PhysRevLett.87.102302,PhysRevC.66.042201}.
However, in contrast to SRCs, they were related mostly to the final state structure of nuclei rather than to its initial state (see, e.g., Ref.\cite{PhysRevLett.84.3045}).\\

@@

is via the exclusive deuteron
electro-disintegration reaction at internal momenta $p_{\mathrm{r}}>300$ MeV/c. For $^{2}\mathrm{H}(e,e'p)n$, within the plane wave impulse approximation (PWIA), the virtual photon couples to
the bound proton which is subsequently ejected from the nucleus without further interaction with the recoiling system (neutron). The neutron carries a recoil momentum, $p_{\mathrm{r}}$, equal in magnitude but opposite in direction
to the initial state proton, $\vec{p}_{\mathrm{r}} \sim -\vec{p}_{\mathrm{i},p}$, thus providing information on the momentum of the bound nucleon and its momentum distribution.\\

@@

\indent The first $^{2}\mathrm{H}(e,e'p)n$ experiments at high $Q^{2}$ ($>1$ (GeV/c)$^{2}$) were carried out at Jefferson Lab (JLab) in Halls A \cite{PhysRevLett.107.262501} and B \cite{PhysRevLett.98.262502}. Both

>>>>

practically unexplored and not well understood. Understanding the high momentum structure of the proton-neutron ($pn$) system is highly important for nuclear physics due to the observed 
dominance of short-range correlations (SRC) in nuclei at nucleon momenta above the fermi momentum. This dominance has been well established by a series of recent experiments
carried out at Jefferson Lab (JLab) and Berkeley National Lab (BNL). In these experiments nucleon momenta up to $p_{\mathrm{r}}\sim1$ GeV/c have been probed and recoil momentum distributions
have been compared to the high momentum part of theoretical deuteron momentum distributions \cite{PhysRevLett.97.162504,PhysRevLett.113.022501,PhysRevLett.122.172502,PhysRevLett.124.212501}.
Recoil momenta up to $p_{\mathrm{r}}\sim1$ GeV/c have also been probed in a $^{3}\mathrm{He}(e,e'p)$ experiment \cite{PhysRevLett.94.082305,PhysRevLett.94.192302} but at a relatively low momentum transfer of 1.5 (GeV/c)$^{2}$ and at a kinematic region
(Bjorken $x_{\mathrm{Bj}}\sim1$) where the cross section is dominated by final state interactions.\\
%===Original introguction (replaced by above intro, suggested by W. Boeglin, in response to referee. comments)====
%At such small inter-nucleon distances, the nucleon-nucleon ($NN$) interaction is expected to become repulsive and the interacting
%nucleons begin to overlap. The dynamics in this short distance region are directly related to the dynamics of two-nucleon short range correlations (SRC) observed in $A>2$ nuclei \cite{PhysRevC.68.014313,PhysRevLett.96.082501,PhysRevLett.99.072501,Fomin_2017,Hen2014,RevModPhys.89.045002}.
%Short-range studies of the deuteron are also important in determining whether, or to what extent, the description of nuclei in terms of nucleon/meson degrees of freedom is still valid, before
%having to include explicit quark degrees of freedom, an issue of fundamental importance in nuclear physics \cite{sargsian_2015}. As of the present time, there are only a few nuclear physics experiments for
%which a transition between nucleonic to quark degrees of freedom has been observed \cite{PhysRevLett.81.4576,PhysRevLett.87.102302,PhysRevC.66.042201}.
%However, in contrast to SRCs, they were related mostly to the final state structure of nuclei rather than to its initial state (see, e.g., Ref.\cite{PhysRevLett.84.3045}).\\

@@

is via the exclusive deuteron electro-disintegration reaction at internal momenta $p_{\mathrm{r}}>300$ MeV/c. For $^{2}\mathrm{H}(e,e'p)n$, within the plane wave impulse approximation (PWIA),
the virtual photon couples to the bound proton which is subsequently ejected from the nucleus without further interaction with the recoiling system (neutron). The neutron carries a recoil
momentum, $p_{\mathrm{r}}$, equal in magnitude but opposite in direction to the initial state proton, $\vec{p}_{\mathrm{r}} \sim -\vec{p}_{\mathrm{i},p}$, thus providing information on the momentum
of the bound nucleon and its momentum distribution.\\

@@

\indent The first $^{2}\mathrm{H}(e,e'p)n$ experiments at high $Q^{2}$ ($>1$ (GeV/c)$^{2}$) were carried out at JLab in Halls A \cite{PhysRevLett.107.262501} and B \cite{PhysRevLett.98.262502}. Both

### 2020-10-07 MY_PUBLICATIONS/PRL/FINAL_DRAFT/PRL_Submission/prl_response/prl_reply/ms.tex
added PRL reply

For each spectrometer, a logic signal was created from  the coincidence of hits in three of the four scintillator planes. The event trigger was the coincidence of these two signals. \\

@@

For the two highest momentum settings ($p_{\mathrm{r}}=580, 750$ MeV/c), a weighted average of the cross sections were taken in the overlapping regions of $p_{\mathrm{r}}$. The results from the previous experiment \cite{PhysRevLett.107.262501} at a $Q^{2}=3.5\pm0.25$ (GeV/c)$^{2}$ are plotted as well (red square). %The good agreement between the Hall A and C data at lower $p_{\mathrm{r}}$ gives us confidence in the measurements made at higher missing momentum settings for which no previous data exist.

>>>>

For each spectrometer, a logic signal was created from  the coincidence of hits in at least three of the four scintillator planes. The event trigger was the coincidence of these two signals. \\

@@

For the two highest momentum settings ($p_{\mathrm{r}}=580, 750$ MeV/c), a weighted average of the reduced cross sections were taken in the overlapping regions of $p_{\mathrm{r}}$. The results from the previous experiment \cite{PhysRevLett.107.262501} at a $Q^{2}=3.5\pm0.25$ (GeV/c)$^{2}$ are plotted as well (red square). %The good agreement between the Hall A and C data at lower $p_{\mathrm{r}}$ gives us confidence in the measurements made at higher missing momentum settings for which no previous data exist.
